---
title: "How to pass parameters to a RMarkdown file?"
author: "Pablo Rodríguez-Sánchez"
date: "4/29/2021"
output: html_document
params:
  who: "world!"
  outputfile: "output.txt"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This is a minimal example about how to pass parameters to a RMarkdown file.
Our purpose is to create a very simple "Hello world" script that does two
things:

1. Says hello.
2. Saves the greetings to a file.

We want to program it in a way that the name of the person to be greeted can be
changed from **outside** the file. For this, we'll need to use parameters.

# Using parameters

## Default parameters

The default parameters are found in the YAML header of this document (first 10 
lines). These parameters will be used if knitting without further details.

For using other parameters, edit the values in the YAML or use `Knit with parameters`
in the drop-down menu in `RStudio`.

```yaml
...
params:
  who: "world!"
  outputfile: "output.txt"
```

## From R console

If the knitting process is launched from the R console, the parameters can be 
overriden by:

```r
rmarkdown::render('parameterized.Rmd', params = list(who = 'Pablo'))
```

## From UNIX console

If the knitting process is launched from the UNIX console, the parameters can be
overriden in a similar way:

```sh
Rscript -e "rmarkdown::render('parameterized.Rmd', params = list(who = 'Pablo'))"
```

Usually, when we are working in the UNIX console we want to use environment 
variables. This can be done by:

```sh
Rscript -e "rmarkdown::render('parameterized.Rmd', params = list(who = Sys.getenv('USER')))"
```

or, with more traditional shell notation:

```sh
Rscript -e "rmarkdown::render('parameterized.Rmd', params = list(who = '${USER}'))"
```

**Note**: the use of " instead of ' as the outermost string marker is relevant in
this last case (i.e.: `Rscript -e 'rmarkdown::render("parameterized.Rmd", params = list(who = "${USER}"))'` will fail).

## From shell script

From what we learned in the previous subsection, using a shell script is 
straightforward.

The contents of the shell script should be something like this:

```sh
#!/bin/sh

# Set parameters
NAME="Pablo"

# Execute
Rscript -e "rmarkdown::render('parameterized.Rmd', params = list(who = '${NAME}'))"
```

After saving, make the script executable by: `chmod +x script_filename.sh`, 
and execute it by: `./script_filename.sh`.

---

# The script

## Parse parameters

After parsing, the parameters' values are available in the rest of the script, 
just as usual. Note that the parameters will be named elements of the list 
`params`.

```{r parse-params}
who <- params$who
outputfile <- params$outputfile
```

## Do something

Greet someone.

```{r create-hello-world}
str <- paste("Hello", who)
print(str)
```

## Create side-effect

In this case, saving a file.

```{r save}
fileConn <- file(outputfile)
writeLines(str, fileConn)
close(fileConn)
```
